---
title: "Basic UI"
author: "Allama Ikbal Sijan"
runtime: shiny
date: "2024-11-29"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Input Functions

All the example codes goes within the `fluidpage()` function

### Free Text Input

```{r, eval=FALSE}
textInput("name", "What's your name?"),
passwordInput("password", "What's your password?"),
textAreaInput("story", "Tell me about yourself", rows = 3)
```

### Numeric Inputs
```{r, eval=FALSE}
numericInput("num", "Number one", value = 0, min = 0, max = 100),
sliderInput("num2", "Number two", value = 50, min = 0, max = 100),
sliderInput("rng", "Range", value = c(10, 20), min = 0, max = 100)
```

### Dates
```{r, eval=FALSE}
dateInput("dob", "When were you born?"),
dateRangeInput("holiday", "When do you want to go on vacation next?")
```

### Limited Choices
```{r, eval=FALSE}
selectInput("state", "What's your favourite state?", state.name),
radioButtons("animal", "What's your favourite animal?", animals),
radioButtons("rb", "Choose one:",
    choiceNames = list(
      icon("angry"),
      icon("smile"),
      icon("sad-tear")
    ),
    choiceValues = list("angry", "happy", "sad")
  ),
selectInput(
    "state", "What's your favourite state?", state.name,
    multiple = TRUE
  ),
checkboxGroupInput("animal", "What animals do you like?", animals),
checkboxInput("cleanup", "Clean up?", value = TRUE),
checkboxInput("shutdown", "Shutdown?")
```

### File Upload
```{r, eval=FALSE}
fileInput("upload", NULL)
```

### Action Button
```{r, eval=FALSE}
actionButton("click", "Click me!"),
actionButton("drink", "Drink me!", icon = icon("cocktail"))

fluidRow(
    actionButton("click", "Click me!", class = "btn-danger"),
    actionButton("drink", "Drink me!", class = "btn-lg btn-success")
  ),
  fluidRow(
    actionButton("eat", "Eat me!", class = "btn-block")
  )
```

## Input Exercise 
### Exercise 

When space is at a premium, it’s useful to label text boxes using a placeholder that appears inside the text entry area. How do you call textInput() to generate the UI below?
```{r}
library(shiny)

ui <- fluidPage(
  textInput("name", "Name", placeholder = "Your name"),
  textOutput("greet")
)

server <- function(input, output, session){
  output$greet <- renderText({
    paste0("Hello ", input$name)
    })
}

shinyApp(ui, server)
```

### Exercise

Carefully read the documentation for sliderInput() to figure out how to create a date slider, as shown below.

```{r}
ui <- fluidPage(
  sliderInput("date", "When should we deliver?", 
              min = as.Date("2020-09-16"),
              max = as.Date("2020-09-23"),
              value = as.Date("2020-09-17"),
              timeFormat = "%F")
)

server <- function(input, output, session){}

shinyApp(ui, server)
```

### Exercise

Create a slider input to select values between 0 and 100 where the interval between each selectable value on the slider is 5. Then, add animation to the input widget so when the user presses play the input widget scrolls through the range automatically.

```{r}
ui <- fluidPage(
  sliderInput("num", "Select a number", min = 0, max = 100, value = 50, step = 5, 
              animate = TRUE)
)
server <- function(input, output, session){}

shinyApp(ui, server)
```

### Exercise

If you have a moderately long list in a selectInput(), it’s useful to create sub-headings that break the list up into pieces. Read the documentation to figure out how. (Hint: the underlying HTML is called <optgroup>.)

```{r}
ui <- fluidPage(
  selectInput("state", "State:",
                  choices = list(`East Coast` = c("NY", "NJ", "CT"),
                                 `West Coast` = c("WA", "OR", "CA"),
                                 `Midwest` = c("MN", "WI", "IA"))),
  textOutput("result")
)
server <- function(input, output, session){
  output$result <- renderText({
    paste("You chose", input$state)
  })
}

shinyApp(ui, server)
```

# Output Functions

### Text
```{r, eval=FALSE}
ui <- fluidPage(
  textOutput("text"),
  verbatimTextOutput("code")
)

server <- function(input, output, session) {
  output$text <- renderText("Hello friend!")
  output$code <- renderPrint(summary(1:10))
}
```

### Tables
```{r}
library(shiny)

ui <- fluidPage(
  tableOutput("static"),
  dataTableOutput("dynamic")
)

server <- function(input, output, session){
  output$static <- renderTable(head(mtcars))
  output$dynamic <- renderDataTable(mtcars, options = list(pagelength = 10))
}
shinyApp(ui, server)
```

### Plots {#plots}
`plotOutput()` has a number of arguments like `click`, `dblclick`, and `hover`. `click = "plot_click"`, they’ll create a reactive input (`input$plot_click`)
```{r}
library(shiny)

ui <- fluidPage(
  plotOutput("plot", width = "400px")
)

server <- function(input, output, session){
  output$plot <- renderPlot(plot(1:10), res = 96)
}

shinyApp(ui, server)
```
### Downloads
```{r, eval=FALSE}
downloadButton()
downloadLink()
```

## Output Exercise

### Exercise

Re-create the Shiny app from Section \@ref(plots) this time setting height to 300px and width to 700px. Set the plot “alt” text so that a visually impaired user can tell that its a scatterplot of five random numbers.

```{r}
library(shiny)

ui <- fluidPage(
  plotOutput("plot", width = "700px", height = "300px")
)

server <- function(input, output, session){
  output$plot <- renderPlot({
    plot(1:5)}, res = 96)
}
shinyApp(ui, server)
```

### Exercise
Update the options in the call to renderDataTable() below so that the data is displayed, but all other controls are suppressed (i.e., remove the search, ordering, and filtering commands). You’ll need to read ?renderDataTable and review the options at https://datatables.net/reference/option/.

```{r}
library(shiny)

ui <- fluidPage(
  dataTableOutput("table")
)

server <- function(input, output, session) {
  output$table <- renderDataTable(mtcars, 
                                  options = list(
                                    searching = FALSE, ordering = FALSE, paging = FALSE))
}
shinyApp(ui, server)
```

### Exercise
Alternatively, read up on reactable, and convert the above app to use it instead.
```{r}

```




